### 2012-07-30 Doc/VRIPHYS2012/vriphys2012.tex
trying new image for membrane/bending deformations
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@436 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b
section 3.1
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@437 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

@@

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
% TODO: anonymize
\author[...]
       {...} % TODO

@@

\noindent
and variation of central point $\delta P_{10}$ \CD{ this formulation would be simplified...} :

@@

In the previous section we present an interpolation based on cubic B\'ezier triangles and a kinematic link between the control points and the element 6DoF Nodes.
In this section we show how this interpolation can be used to compute a finite element formulation of the shell equations.

@@

\begin{equation}
\left[ \begin{array}[c]{c} \xi_1(x,y) \\ \xi_2(x,y) \\ \xi_3(x,y) \end{array} \right] =
\left[ \begin{array}[c]{ccc}
1 & 1 &  1 \\
x_1^0 & x_2^0 & x_3^0 \\
y_1^0 & y_2^0 & y_3^0
\end{array} \right] ^{-1} 
\left[ \begin{array}[c]{c} 1 \\ x \\ y \end{array} \right] 
\end{equation}

Thus, we can locally compute some spacial derivatives, along $x$ and $y$ on the interpolation function vector $T$ (which has three components $(T^x, T^y, T^z)$ .
The 2D tensor $F$ of the in-plane deformations can be computed as a $2\times2$ matrix where the line $k$ and the column $l$ alternatively represent $x$ and $y$\footnote{It can be noted $k=\{x,y\}$ and $l=\{x,y\}$}:
\begin{equation}
F_{kl} = \frac{\delta T^k}{\delta l}  =  \sum_{j=1}^3 \frac{\delta T^k}{\delta \xi_i} \frac{\delta \xi_i}{\delta l}  
\end{equation}

For the shell formulation, there is no need to compute the derivatives along direction $z$, but the second derivatives derivatives $\frac{\delta T^z}{\delta kl}$...

@@

Shell element is a combination of two simpler elements describing two types
of deformations:

@@

    \item Elastic membrane defining deformations in plane of the
    element. It encompasses deformations like stretching and shearing and
    is described by 3 degrees of freedom: in-plane displacements $u_x, u_y$
    and rotation $\theta_z$ around the axis perpendicular to the plane of
    the element (Figure \ref{fig-membrane}).

\Figure{0.8\linewidth}{membrane}
{DOFs of elastic membrane element.}

    \item Bending plate defining out of plane bending deformation. It is
    described by 3 degrees of freedom: rotation around two axes $\theta_x,
    \theta_z$ and out of plane displacement $u_z$ (Figure \ref{fig-plate}).

@@

\Figure{0.8\linewidth}{plate}
{DOFs of bending plate element.}

@@

\subsection{Mechanics} % {{{

@@

>>>>

% For large/wide figures
%\FigureW{width}{name}{caption}
\newcommand{\FigureWI}[3]{%
\begin{figure*}[htb]
  \centering
  \input{#1.pdf_tex}
  \caption{\label{fig-#1}#2}
\end{figure*}}

@@

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
% TODO: anonymize
\author[Authors]
       {Authors} % TODO

@@

\noindent
and variation of central point $\delta P_{10}$ \CD{ this formulation would be modified...} :

@@

In the previous section we present an interpolation based on cubic B\'ezier triangles and a kinematic link between the control points and the element 6DoF Nodes.
In this section we show how this interpolation can be used to compute a finite element formulation of the shell equations.
The model relies on large transformation formulation: the geometrical non-linearities are handled using a corotational formulation, whereas, in the frame of the element, infinitesimal strain theory is used.

@@

\begin{equation}
\label{shape-vector}
\left[ \begin{array}[c]{c} \xi_1(x,y) \\ \xi_2(x,y) \\ \xi_3(x,y) \end{array} \right] =
\left[ \begin{array}[c]{ccc}
1 & 1 &  1 \\
x_1^0 & x_2^0 & x_3^0 \\
y_1^0 & y_2^0 & y_3^0
\end{array} \right] ^{-1} 
\left[ \begin{array}[c]{c} 1 \\ x \\ y \end{array} \right] 
\end{equation}

Thus, we can locally compute some spacial derivatives, along $x$ and $y$ on the interpolation function vector $T$ (which has three components $(T^x, T^y, T^z)$ .
The 2D tensor $F$ of the in-plane deformations can be computed as a $2\times2$ matrix where the line $k$ and the column $l$ alternatively represent $x$ and $y$\footnote{It can be noted $k=\{x,y\}$ and $l=\{x,y\}$}:
\begin{equation}
F_{kl} = \frac{\partial T^k}{\partial l}  =  \sum_{i=1}^3 \frac{\partial T^k}{\partial \xi_i} \frac{\partial \xi_i}{\partial l}  
\end{equation}

The term $\frac{\partial T^k}{\partial \xi_i}$ can be derived from the expression of $T(\xi_1,\xi_2, \xi_3)$ in equation (\ref{global-transform}) and 
$\frac{\partial \xi_i}{\partial l}$ from equation (\ref{shape-vector}). 
The in-plane displacement gradient can be obtained with tensor $F$:
\begin{equation}
\frac{\partial u^k}{\partial l} = F_{kl} - I_{kl} 
\end{equation}
where $I$ is the identity matrix ($I_{kl}  = \{1$ if $l=k$ ;  $0$ if $l \neq k \} $). 
The in-plane displacement gradient can be linked to the degrees of freedom of the shell using the expression of the displacement of a point $\delta T$ (see equation  (\ref{bez-delta})) on the shell surface:
$$
\frac{\partial u^k}{\partial l} = \frac{\partial (\delta T^k)}{\partial l}
$$

In the local frame of the element, we rely on infinitesimal strain theory, thus the expression of the strain tensor (using Voigt notations) is:
\begin{equation}
\label{in-plane-strain}
\left[ \begin{array}[c]{c} \epsilon_{xx} \\ \epsilon_{yy}  \\ 2\epsilon_{yx} \end{array} \right] = 
\left[ \begin{array}[c]{c} \frac{\partial u^x}{\partial x}  \vspace{1mm}\\ \frac{\partial u^y}{\partial y}   \\ \frac{\partial u^x}{\partial y}  + \frac{\partial u^y}{\partial x}  \end{array} \right]
\end{equation}

For the bending formulation, the shell theory uses a measure of the out-of-plane-displacement $u^z(x,y)$. 
The strain tensor has additional contributions based on Kirchhoff-Love theory for thin plates:
\begin{equation}
\label{out-plane-strain}
\left[ \begin{array}[c]{c} \epsilon_{xx} \\ \epsilon_{yy}  \\ 2\epsilon_{yx} \end{array} \right] +=
\left[ \begin{array}[c]{c} -z \frac{\partial^2 u^z}{\partial x^2}  \vspace{1mm}\\ -z \frac{\partial^2 u^z}{\partial y^2}   \vspace{1mm}\\ - 2z \frac{\partial^2 u^z}{\partial xy}    \end{array} \right]
\end{equation}
where $z$ is the local coordinate along $z$-axis of the point on which the strain is measured.
\CD{This is something that is maybe wrong in our formulation... as the bending adds a contribution on the deformation tensor, maybe we should not separate the contribution from membrane and bending as we do now ($K_m$ and $K_b$ are computed separately) }

As the term  $\frac{\partial \xi_i}{\partial l}$ provides a constant value, the second derivatives can be computed quite easily, for instance, for $\epsilon_{xx}$:
\begin{equation}
\frac{\partial^2 u^z}{\partial x^2}  = \sum_{i=1}^3\frac{\partial}{\partial \xi_i}\left(\sum_{j=1}^3 \frac{\partial T^z}{\partial \xi_j} \frac{\partial \xi_j}{\partial x}\right) \frac{\partial \xi_i}{\partial x} 
=  \sum_{i=1}^3  \sum_{j=1}^3 \frac{\partial^2 T^z}{\partial \xi_i\xi_j}   \frac{\partial \xi_j}{\partial x} \frac{\partial \xi_i}{\partial x}  
\end{equation}

@@

As previously describes, the shell element's strain is a combination of two types of deformations, as shown in figure \ref{deformations}.
If we assume that the initial out-plane deflections are negligible and that the frame of the shell element nodes are aligned with local $z$ axis, we can distribute these two deformations into two groups of degrees of freedom at the nodes' level.

\FigureWI{deformations}
{Plate deformations:%
a) membrane deformations like stretching or shearing, due to in-plane displacements
b) undeformed;
c) bending deformation, due to out-plane displacements
\label{deformations}}

@@

    \item Elastic membrane defining deformations in plane of the
    element. It encompasses deformations like stretching and shearing.
    It corresponds to the in-plane displacements $u_x, u_y$. 
    At the nodes' level, it corresponds to in-plane displacements $\delta\mathbf{x}_x, \delta\mathbf{x}_y$ and rotation $\theta_z$.
(Figure \ref{fig-membrane}).
    \item Bending plate defining deformation due to out of plane bending displacement $u_z$. It is
    described by 3 degrees of freedom: rotation around two axes $\theta_x,
    \theta_z$ and out of plane displacement $\delta\mathbf{x}_z$ (Figure \ref{fig-plate}).

@@

@@

\subsection{Stiffness matrix} % {{{

In order to compute FEM-based stiffness for the shell element, we need to compute the jacobian matrix between strains and nodes' displacements.
For that we can reuse equation (\ref{bez-delta}) and apply the derivatives. For instance, for the in-plane displacement:
$$
\frac{\delta \epsilon_{xx}} {\delta \mathbf{x}_{2}} = \sum_{i=1}^{3} (\frac{\partial n_2}{\xi_i} + \frac{\partial n_6}{\xi_i} + \frac{\partial n_9}{\xi_i} + \frac{1}{3}\frac{\partial n_{10}}{\xi_i})\frac{\partial \xi_i}{\partial x}  
$$

@@

\CD{or... $K =  \iiint_V \left[\begin{matrix} J_m & J_b \end{matrix} \right]^T  M \left[\begin{matrix} J_m & J_b \end{matrix} \right]$. Could we test that option easily in the code, or is it a nightmare ??}
